---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ----- comment line-by-line ----- #

# The line below... creates a sequence from 0-8 by one increment to the time in years.
time_yr <- seq(from = 0, to = 8, by = 1)

# The line below... creates a 9x4 matrix where fish_year0 is the length of the column and time_yr is the length of the row
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# The line below...maps the proj_year matrix to fish_year0, populates the proj_yr matrix where the right ROW (row 1) will be populated by the initial condition of fish_yr0. only populates first row
proj_year[1, ] <- fish_year0

# The for loop below... loops the row starting at 2 where the projection year is subtracted by 1 with each iteration. seq_along is not used bc row 1 has already been populated. starting in row two, this takes the dot product of the matrix and the row containing the vector from the previous cycle.
for (i in 2:length(time_yr)) {
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}


# The rest is wrangling & visualization (run to check it out):
#this reasigns the column names of proj_yr since they are unnamed
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

#proj_df binds together the vector of times with matrix of population projections
#pivot_longer melts into long format from wide format. where life stage is split over 4 different columns and turns it into one column. the second column stage_population takes the values in those columns and aligns them next to each other appropriately. -time_yr excludes time_yr
proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

#because of the pivot_longer, it only requires one geom_line to graph all different lifestages
ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 


